Exploring probability distributions for bivariate temporal granularities

### 2019-11-25 main.Rmd
arrangment section fine tune

According to this definition, $G \trianglelefteq H,$ if each granule H (j) is the union of some
granules of G. For example, $Days \trianglelefteq Months,$. However, this  relationship is not fully described until we associate it with periodicity. For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping. If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year. Considering leap years and all their exceptions, the period becomes 400 years. 
\begin{timedom}\label{def:linear}
A granularity H is periodical with respect to a granularity G if
(1) $G \trianglelefteq H,$, and
(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
$H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{timedom}

@@

**Definition: Circular granularity** A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if

\begin{equation} \label{eq:eq2}
C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
BG(z) \neq \phi
\end{equation}
where,  
z denotes the index set.  
BG denotes the index of the tsibble (bottom granularity).  
G be a linear granularity, periodic with respect to BG.    
k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.

**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram \autoref{fig:circular-gran}. Each granule is represented by a box. The diagram also illustrate that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and the bottom granularity. The index of the tsibble considered is days. G and H are two linear granularities denoting days and weeks respectively. The period length of days with respect to weeks is 7, hence, $C_{G, BG}(z)$ represented by C(BG, G) is given by $BG(z\mod 7)$. The circular granularity C(BG, G) consists of repetitive pattern {BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)} which repeats itself after each period length. Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14. C(BG, H) also repeats the pattern ${BG(0), BG(1), \dots, BG(13)}$ in every period. Hence, circular granularities are represented as repetitive categorization of time in the diagram.

@@

\begin{equation}\label{eq:eq3}
A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert) for z \in T_k
\end{equation}
where,  
z denotes the index set (index of BG),
R' denotes the number of granules of AG in each period of BG,  
$T_w$ are the sets of indices of BG describing $AG_w$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
$\vert T_w \vert$ is the cardinality of set $T_w$.

@@

**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:aperiodic-gran}. Each granule is represented by a box. Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity. 
Two linear granularities AG and AH are considered, where $AG = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$. This implies that AG is made up of shrinking every 1st granule of Group_3(BG) by 1 and AH is made up of shrinking every 1st granule of Group_3(BG) by 2. Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different. A(BG, AG) and A(BG, AH) are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH. For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}. For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.

@@

The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity. The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG. It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (\autoref{def:quasi})

@@

for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T(w)$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.

>>>>

According to this definition, $G \trianglelefteq H,$ if each granule H (j) is the union of some
granules of G. For example, $Days \trianglelefteq Months$. However, this  relationship is not fully described until we associate it with periodicity. For example, $Days \trianglelefteq Months,$, since each month is a grouping of a number of days, however it is also a periodical grouping. If leap year is ignored, the periodicity of the grouping is 1 year, since each month would be defined as the grouping of the same number of days (31, 28, or 30, depending on the month) every year. Considering leap years and all their exceptions, the period becomes 400 years. 
\begin{timedom}\label{def:linear}
A granularity H is periodical with respect to a granularity G if
(1) $G \trianglelefteq H,$, and
(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
$H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{timedom}

@@

\begin{timedom}\label{def:circular}
A circular granularity C(BG, G) relates a linear granularity G to the bottom granularity, if

\begin{equation} \label{eq:eq2}
C_{BG, G}(z) = BG(z\mod k(BG,G)) \forall z \in \mathbb{Z}^+,
BG(z) \neq \phi
\end{equation}
where,  
z denotes the index set.  
BG denotes the index of the tsibble (bottom granularity).  
G be a linear granularity, periodic with respect to BG.    
k(BG,G) be the period length of the bottom granularity with respect to linear granularity G.  
\end{timedom}

**Discussion**: Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}. Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and the bottom granularity. The index of the tsibble considered is days. G and H are two linear granularities denoting days and weeks respectively. The period length of days with respect to weeks is 7, hence, $C_{G, BG}(z)$ represented by C(BG, G) is given by $BG(z\mod 7)$. The circular granularity C(BG, G) consists of repetitive pattern {BG(0), BG(1), BG(2), BG(3), BG(4), BG(5), BG(6)} which repeats itself after each period length. Similarly, C(BG, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14. C(BG, H) also repeats the pattern ${BG(0), BG(1), \dots, BG(13)}$ in every period. Hence, circular granularities are represented as repetitive categorization of time in the diagram.

@@

\begin{equation}\label{eq:eq3}
A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert), z \in T_k
\end{equation}
where,  
z denotes the index set (index of BG),
R' denotes the number of granules of AG in each period of BG,  
$T_w$ are the sets of indices of BG describing $AG(w)$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$, and $T_w = 0$ for $w \epsilon Z_{<0}$  
$\vert T_w \vert$ is the cardinality of set $T_w$.

@@

**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:aperiodic-gran}. Each granule is represented by a box. Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity. 
Two linear granularities AG and AH are considered, where $AG = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$. This implies that AG is made up by shrinking every 1st granule of Group_3(BG) by 1 granule and AH is made up of shrinking every 1st granule of Group_3(BG) by 2 granules. Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different. A(BG, AG) and A(BG, AH) are repetitive categorization of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH. For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}. For AH, R' = 2, P'= 12, $T_0$ = {0, 1, 2, 3, 4, 5, 6} and $T_1$ = {7, 8, 9, 10, 11}, then we will have \autoref{eq:eq4} and  \autoref{eq:eq5}.

@@

The linear granularities G, H, AG and AH considered in the definitions of circular and aperiodic granularities are all periodic with respect to the bottom granularity. The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG. It can be noted here that if a linear granularity AG is quasi-periodic with respect to BG, then \autoref{eq:eq3} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi})

@@

for $z \in T_k$ and $e_k$ is the number of exceptions in $\bigcup\limits_{w = 0}^{k} T_w$ and $E_{u} = [e_{begin}(u), e_{end}(u)]$.
